
\subsection{Software Lifecycle}
\label{sec:software-lifecycle}

The two most important goals of the CENELEC EN 50128 standard are the separation
of the lifecycle into well-defined phases and to produce and record all
documentation of the development process. In addition, the standard specifies
some important constrains which must be fulfilled by a development
process. These constraints mainly result from the application in safety critical
domains.

Naturally, the document-centric nature and the additional constraints are not
necessarily fulfilled by generic software development processes which have been
developed for non-critical domains. In particular for agile development
processes like SCRUM, an adaptation in some aspects will be
required. In openETCS the development process will be less  focused on
extracting the real wishes from the client from user-stories, but on finding the
intended semantics of the ERTMS specification.

\subsection{Organisational Structure and Roles}
\label{sec:organ-struct-roles}

CENELEC EN 50128 defines a set of required roles for the development of SIL3 or
SIL4 compliant software (e.g. requirements manager, assessor, etc., for a full
list see 5.1 of CENELEC EN 50128). Theses roles and which of these roles can be
filled by the same person are specified in 5.1.2.10 of the standard. This
distribution and in particular the required separation of several of the roles
will be mandatory for the development process.

\subsection{Consistency of Documentation}
\label{sec:cons-docum}

For the development according to CENELEC EN 50128, it is mandatory to have a
consistent definition and usage of all terms, abbreviations etc. This is
especially important if the development process uses small scale iterations, as
the consistency must be assured and restores in necessary after each
iteration. In particular, it must be assured at all times, that dependent
documents, i.e., in hierarchical relation do not contradict each other.

\subsection{Traceability of Requirements}
\label{sec:trac-requ}

For each of the generated documents, the traceability must be assured by
referencing all other concerned documents and by specifying unique reference
numbers. Just as consistency, this traceability is a very important aspect which
is often not a concern in more general software development processes.



%% some ideas of agile in safety critical development:
%% see open-DO project, presentation Peter Gardner from AdaCore
%% www.open-do.org
%%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "wp-2.2"
%%% End:
